The `cheering3()` probability model for order statistics: a

### 2023-01-05 inst/multiset_hyper2.Rmd
clarifications and rewording

---
title: "multiset_hyper2"
output: html_document
date: "2022-07-28"
---

@@

```{r defineM}
M <- apply(multiset(c(1,1,2,2,3)),2,function(x){letters[x]})
colnames(M) <- rep("",ncol(M))   # tidier and more compact appearance
M <- noquote(M)
M
```

Above, we use `multiset()` to enumerate the sample space of five
competitors comprising two twins denoted `a`, two twins denoted `b`,
and a singleton `c`.  This document discusses the issue of whether the
twins are identical twins and hence indistinguishable.  Matrix `M` has
${5\choose2\,2\,1}=30$ columns, one per possible result.  We will
calculate the likelihood of each column and calculate their sum:

@@

To explain the factor of $1/4$ discussed above we need some new
notation.  We have five competitors who I will call $\mathbf{do}$,
$\mathbf{re}$, $\mathbf{me}$, $\mathbf{fa}$, and $\mathbf{sol}$.  Now
$p_X$ means "the Bradley Terry strength of $X$''.  The setup here is
that $\mathbf{do}$ and $\mathbf{re}$ are twins, and that are twins
(and $\mathbf{sol}$ is a singleton).  We have that the Bradley-Terry
stregth of $\mathbf{do}$ and $\mathbf{re}$ is `a`, the strength of
$\mathbf{me}$ and $\mathbf{fa}$ is `b`, and the strength of
$\mathbf{sol}$ is `c`.

Consider, for example, the first column of `M` which is `a c b b a`.
In theory, this means the competitor who came first had strength `a`,
the competitor who came second had strength `c`.  So the order
statistic is one of

@@

heAbove we see that $\mathbf{do}$ and $\mathbf{re}$ come first and last
(or last and first), and $\mathbf{me}$ and $\mathbf{fa}$ come third
and fourth (or fourth and third), while $\mathbf{sol}$ comes second.
All four possibilities have the same probability and all four are
admissible order statistics for the five solfeggio syllables.

>>>>

---
title: "The `cheering3()` probability model for order statistics: a
verification with the `multiset()` function of the `partitions` package"
output: html_document
date: "2022-07-28"
---

@@

```{r defineM}
M <- apply(partitions::multiset(c(1,1,2,2,3)),2,function(x){letters[x]})
colnames(M) <- rep("",ncol(M))   # tidier and more compact appearance
M <- noquote(M)
M
```

Above, we use `multiset()` to enumerate the sample space of five
competitors comprising two twins denoted `a`, two twins denoted `b`,
and a singleton `c`.  This document discusses the issue of whether the
twins are identical twins and hence indistinguishable.  Matrix `M` has
${5\choose2\,2\,1}=30$ columns, one per possible result.  We will
calculate the likelihood of each column and sum them to get the
probability of observing any one of the 30 order statistics [the
observations are mutually exclusive, so
$\operatorname{Prob}\left(x_1\cup\cdots\cup x_n\right) =
\operatorname{Prob}\left(x_1\right)+\cdots
+\operatorname{Prob}\left(x_n\right)$]:

@@

To explain the factor of $1/4$ discussed above we need some new
notation.  We have five competitors who I will call $\mathbf{do}$,
$\mathbf{re}$, $\mathbf{me}$, $\mathbf{fa}$, and $\mathbf{sol}$.  Now
$p_X$ means "the Bradley Terry strength of $X$''.  The setup here is
that $\mathbf{do}$ and $\mathbf{re}$ are twins, and that $\mathbf{me}$
and $\mathbf{fa}$ are twins (and $\mathbf{sol}$ is a singleton).  We
have that the Bradley-Terry strength of $\mathbf{do}$ and
$\mathbf{re}$ is `a`, the strength of $\mathbf{me}$ and $\mathbf{fa}$
is `b`, and the strength of $\mathbf{sol}$ is `c`.

Consider, for example, the column of `M` reading `a c b b a`.  In
theory, this means the competitor who came first had strength `a`, the
competitor who came second had strength `c`.  So the order statistic
is one of the $2!2!1!=4$ following orders:

@@

Above we see that $\mathbf{do}$ and $\mathbf{re}$ come first and last
(or last and first), and $\mathbf{me}$ and $\mathbf{fa}$ come third
and fourth (or fourth and third), while $\mathbf{sol}$ comes second.
All four possibilities have the same probability and all four are
admissible order statistics for the five solfeggio syllables.
